---
title: "Carpentries Rmd Example"
output: github_document
---

# Script purposes and organization
This is an example Rmd script developed during a Carpentries workshop. In this script, we will develop a workflow and report that does the following tasks:
 1. Prepares the R environment
 2. Download and save USGS stream stage data
 3. Tidy, QAQC, and gap-fill the data and save the processed data
 4. Write up some summary statistics.

## 1. Prepare the R environment
We will load any relevant packages and set the default for our code to appear in our final markdown document.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(dataRetrieval)
library(dplyr)
library(zoo)
library(lubridate)
library(ggplot2)
```

## 2. Download and save USGS stream stage data.
We will use the `readNWISdv` function from the `dataRetrieval` package to download stream stage data for the USGS gage on the Arkansas River near Larned KS. To do this, we will need to provide the 8-digit gage ID number (07141220), the start/end date, and the parameter code for stage (00065; discharge is 00060).

```{r download-data}

# gage ID
usgs_id <- "07141220"

# download data
data_raw <- 
  readNWISdv(siteNumbers = usgs_id,
             parameterCd = "00065",
             startDate = "2018-10-01",
             endDate = "2021-09-30")

# inspect data
summary(data_raw)

# save data
write.csv(data_raw, "data/ExampleStreamStage_Raw.csv")
```

## 3. Tidy, QAQC, and gap-fill the data and save the processed data

In this section, we will perform some basic data cleaning operations to get our data ready for further analysis.

```{r clean-data}
# create a new data frame with better column names
data_tidy <-
  data_raw %>% 
  rename(stage_ft = X_00065_00003,
         stage_QAcode = X_00065_00003_cd) %>% 
  select(-agency_cd, -site_no)

# look at our new data frame
head(data_tidy)

# first thing to do is always plot!
ggplot(data_tidy, aes(x = Date, y = stage_ft)) +
  geom_line()

# check for missing dates by comparing all possible dates to the dates you have
first_date <- min(data_tidy$Date)
last_date <- max(data_tidy$Date)
all_dates <- seq(first_date, last_date, by = "day")  # make vector of all dates
length(all_dates) == length(data_tidy$Date)  # check for missing dates
missing_dates <- all_dates[!(all_dates %in% data_tidy$Date)]

# add missing dates to your data frame - we will need to make a new df with the same columns, bind them together, and arrange by date
new_dates <- data.frame(Date = missing_dates, 
                        stage_ft = NA,
                        stage_QAcode = "Gapfill")
data_clean <- 
  bind_rows(data_tidy, new_dates) %>% 
  arrange(Date)

summary(data_clean)

# fill in those gaps using linear interpolation, from the zoo package
# ?na.approx
# na.approx(c(1, 2, NA, 4))
data_clean$stage_ft <- na.approx(data_clean$stage_ft) # can set a maxgap if you want

summary(data_clean)

# plot and inspect
ggplot(data_clean, aes(x = Date, y = stage_ft, color = stage_QAcode)) +
  geom_point()

# save data
write.csv(data_raw, "data/ExampleStreamStage_Clean.csv")

```

## 4. Write up some summary statistics

We analyzed data from USGS gage `r usgs_id`. During the period of record, the maximum stage was `r max(data_clean$stage_ft)` and the minimum stage was `r min(data_clean$stage_ft)`.